Predicting _C. difficile_ Infection Severity from the Taxonomic Composition of the Gut Microbiome

### 2023-05-18 paper/paper.qmd
Use 'training set' and 'test set'

Report median AUROC for trainset and testset, and median AUBPRC for testset (@fig-performance).

@@

All sequences were processed with mothur (v1.43) using the MiSeq SOP protocol
[@schloss_introducing_2009; @kozich_development_2013].
Paired sequencing reads were combined and
aligned with the SILVA (v132) reference database [@quast_silva_2013] and taxonomy was
assigned with a modified version of the Ribosomal Database Project reference
sequences (v16) [@cole_ribosomal_2014].
Samples were rarefied to 5,000 sequences per sample, repeated 1,000
times for alpha and beta diversity analysis.

@@

mikropml R package [@topcuoglu_mikropml_2021]

Balanced precision

@@

The workflow was defined with Snakemake [@koster_snakemake_2012] using a custom
version of the mikropml Snakemake workflow [@sovacool_mikropml_2023].
Dependencies were managed with conda environments.
Scripts were written in R [@r_core_team_r_2020],
Python [@van_rossum_python_2009],
and GNU bash.
In addition to the software already cited above, other packages used in the
creation of this manuscript include
cowplot [@wilke_cowplot_2020],
ggtext [@wilke_ggtext_2020],
ggsankey [@sjoberg_ggsankey_2022],
schtools [@sovacool_schtools_2022],
the tidyverse metapackage [@wickham_welcome_2019],
Quarto, and
vegan [@oksanen_vegan_2023].

@@

::: {#fig-performance fig-cap="**Performance of ML models.**
Area Under the Receiver-Operator Characteristic Curve (AUROC) for the testsets
and cross-validation folds of the trainsets,
and the Area Under the Balanced Precision-Recall Curve (AUPRC) for the testsets.
Left: models were trained on the full dataset, with different numbers of samples available for each severity definition.
Right: models were trained on the intersection of samples with all labels available for each definition.
"}
![](figures/ml-performance.png)
:::

>>>>

Report median AUROC for training set and test set, and median AUBPRC for test set (@fig-performance).

@@

All sequences were processed with mothur (v1.46) using the MiSeq SOP protocol
[@schloss_introducing_2009; @kozich_development_2013].
Paired sequencing reads were combined and
aligned with the SILVA (v132) reference database [@quast_silva_2013] and taxonomy was
assigned with a modified version of the Ribosomal Database Project reference
sequences (v16) [@cole_ribosomal_2014].
Sequences were clustered into _de novo_ OTUs with the OptiClust algorithm in
mothur [@westcott_opticlust_2017].
Samples were rarefied to 5,000 sequences per sample, repeated 1,000
times for alpha and beta diversity analysis.
<!-- TODO supplementary figure with alpha and beta diversity & significance -->

@@

TODO mikropml R package [@topcuoglu_mikropml_2021]

Balanced precision TODO

@@

The workflow was defined with Snakemake [@koster_snakemake_2012] using a custom
version of the mikropml Snakemake workflow [@sovacool_mikropml_2023].
Dependencies were managed with conda environments.
Scripts were written in R [@r_core_team_r_2020],
Python [@van_rossum_python_2009],
and GNU bash.
Additional software and packages used in the creation of this manuscript include
cowplot [@wilke_cowplot_2020],
ggtext [@wilke_ggtext_2020],
ggsankey [@sjoberg_ggsankey_2022],
schtools [@sovacool_schtools_2022],
the tidyverse metapackage [@wickham_welcome_2019],
Quarto, and
vegan [@oksanen_vegan_2023].

@@

::: {#fig-performance fig-cap="**Performance of ML models.**
Area Under the Receiver-Operator Characteristic Curve (AUROC) for the test sets
and cross-validation folds of the training sets,
and the Area Under the Balanced Precision-Recall Curve (AUPRC) for the test sets.
Left: models were trained on the full dataset, with different numbers of samples available for each severity definition.
Right: models were trained on the intersection of samples with all labels available for each definition.
"}
![](figures/ml-performance.png)
:::
